Constraining the High Density Nuclear Symmetry Energy with Pions

### 2020-06-14 chapters/ch_data_analysis.tex
Backup

\begin{figure}[H]
\includegraphics[width=\linewidth]{highlowcal.png}
\caption{Calibration of low and high gain sections of the anode wires.}
\label{fig:highlowcal}
\end{figure}

@@

 \begin{table}[!htp] % not just 'h!'
\centering % not a center environment
\begin{tabular}{
  @{}
  l
  S[table-format=1.2]
  S[table-format=1.2]
  S[table-format=1.2]
  S[table-format=5.2]
  S[table-format=5.2]
  @{}
}
\toprule
Beam Energy Loss  &
 {${}^{132}$Sn} &
 {${}^{124}$Sn} &
 {${}^{112}$Sn} &
 {${}^{108}$Sn} &
  {Avg.}\\

@@

\begin{figure}[!htb]
\includegraphics[width=\linewidth]{spacechg_cartoon.png}
\caption{Location of space charge in 132 Sn}
\label{fig:spacechg_cartoon}
\end{figure}

\begin{figure}[!htb]
\includegraphics[width=\linewidth]{beampath.png}
\caption{Beam path of the experiments}
\label{fig:beampaths}
\end{figure}

@@

\begin{figure}[!htb]
\includegraphics[width=\linewidth]{Effect_SC.pdf}
\caption{Example of the map of the electron shift. The effect is also shown on left-going (blue) and right-going (green) tracks. The original tracks are shown in the solid line and the shifted track in the dotted line.}
\label{fig:sc_shift}
\end{figure}

@@

\begin{figure}[!htb]
\includegraphics[width=\linewidth]{DVTP_raw.pdf}
\caption{$\Delta\mathrm{V}_\mathrm{x}$ distribution for left-going and right-going tracks. }
\label{fig:VLR}
\end{figure}

\begin{figure}[!htb]
\includegraphics[width=\linewidth]{VLR_beamrate.png}
\caption{ $\Delta V_{LR}$ versus the beam rate for all systems with the fitted function.}
\label{fig:vlr_br}
\end{figure}

\begin{figure}[!htb]
\includegraphics[width=\linewidth]{sc_beamrate.png}
\caption{Space charge fit function versus the beam rate for all systems. These are the assumed functions we use for interpolating the space charge value from the beam rate.}
\label{fig:spacechg_br_all}
\end{figure}

@@

%\begin{sidewaysfigure}[H]
%\centering
%\includegraphics[width=\textwidth]{FlowEmbedding.pdf}
%\caption{Flow of embedding implementation in the software.}
%\label{fig:flow}
%\end{sidewaysfigure}

%\clearpage
%\pagestyle{lscape} % first clear the page and change the pagestyle
%\begin{landscape}
% your landscape table(s) or figure(s) here
%CONSIDER ROTATING THIS FIGURE IF ALLOWED
\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{FlowEmbedding.pdf}
\caption{Flow diagram of the embedding software.}
\label{fig:flow}
\end{figure}

@@

\section{PID Analysis}
\label{sec:pid}
%PID lines over all
%PID of pion cut
%background estimation

>>>>

\begin{figure}[!htb]
\includegraphics[width=\linewidth]{highlowcal.png}
\caption{Calibration of low and high gain sections of the anode wires.}
\label{fig:highlowcal}
\end{figure}

@@

 \begin{table}[!htb] % not just 'h!'
\centering % not a center environment
\begin{tabular}{
  @{}
  l
  S[table-format=1.2]
  S[table-format=1.2]
  S[table-format=1.2]
  S[table-format=5.2]
  S[table-format=5.2]
  @{}
}
\toprule
Beam Energy Loss  &
 {${}^{132}$Sn} &
 {${}^{124}$Sn} &
 {${}^{112}$Sn} &
 {${}^{108}$Sn} &
  {Avg.}\\

@@

\begin{figure}[!htb]
\centering
\includegraphics[width=\linewidth]{spacechg_cartoon.png}
\caption{Location of space charge in 132 Sn}
\label{fig:spacechg_cartoon}
\end{figure}

\begin{figure}[!htb]
\centering
\includegraphics[width=\linewidth]{beampath.png}
\caption{Beam path of the experiments}
\label{fig:beampaths}
\end{figure}

@@

\begin{figure}[!htb]
\centering
\includegraphics[scale=.5]{Effect_SC.pdf}
\caption{Example of the map of the electron shift. The effect is also shown on left-going (blue) and right-going (green) tracks. The original tracks are shown in the solid line and the shifted track in the dotted line.}
\label{fig:sc_shift}
\end{figure}

@@

\begin{figure}[!htb]
\centering
\includegraphics[scale=.5]{DVTP_raw.pdf}
\caption{$\Delta\mathrm{V}_\mathrm{x}$ distribution for left-going and right-going tracks. }
\label{fig:VLR}
\end{figure}

\begin{figure}[!htb]
\centering
\includegraphics[width=\linewidth]{VLR_beamrate.png}
\caption{ $\Delta V_{LR}$ versus the beam rate for all systems with the fitted function.}
\label{fig:vlr_br}
\end{figure}

\begin{figure}[!htb]
\centering
\includegraphics[width=\linewidth]{sc_beamrate.png}
\caption{Space charge fit function versus the beam rate for all systems. These are the assumed functions we use for interpolating the space charge value from the beam rate.}
\label{fig:spacechg_br_all}
\end{figure}

@@

%\begin{sidewaysfigure}[!htb]
%\centering
%\includegraphics[width=\textwidth]{FlowEmbedding.pdf}
%\caption{Flow of embedding implementation in the software.}
%\label{fig:flow}
%\end{sidewaysfigure}

%\clearpage
%\pagestyle{lscape} % first clear the page and change the pagestyle
%\begin{landscape}
% your landscape table(s) or figure(s) here
%CONSIDER ROTATING THIS FIGURE IF ALLOWED
\begin{figure}[!htb]
\centering
\includegraphics[width=\textwidth]{FlowEmbedding.pdf}
\caption{Flow diagram of the embedding software.}
\label{fig:flow}
\end{figure}

@@

\section{PID Analysis}
\label{sec:pid}
%PID lines over all picture

%PID of pion cut
%background estimation

The energy loss curve in Eq.~\ref{eq:bb} can be described as a 5-parameter general function as,

\begin{equation}
\frac{dE}{dx} = \frac{p_0}{\beta^{p_3}}(p_1 - \beta^{p_3} + \ln(p_2 + {\beta\gamma}^{-p_4})),
\label{eq:5par_bb}
\end{equation}

where the parameters $p_0-p_4$ are free parameters, $\beta= p/\sqrt{p^2 + m^2c^4}$, where $m$ is the mass of the particle and $p$ is the momentum. The PID lines vary slightly as a function of the emission angle of each track. The PID was subdivided into 6 pitch angles, $\theta_P = \arctan(p_y/p_z)$, and 6 yaw angles, $\theta_Y = \arctan(p_x/p_z)$, ranging from \ang{-90} to \ang{90} for both. The pion spectra of each yaw-pitch bin was fit with Eq.~\ref{eq:5par_bb} to get the PID line which best describes the data. The distribution around this mean value is not a Gaussian distribution. The variable $z$ has been used before to transform the energy loss of a particular track $dE/dx$ into a more Guassian variable, defined as

\begin{equation}
z_i = \ln\Big(\frac{\langle dE/dx\rangle}{\langle dE/dx\rangle_i}\Big),
\end{equation}

where $\langle dE/dx\rangle_i$ is the mean energy loss curve fit from Eq.~\ref{eq:5par_bb} for a given particle type $i$. The $z_i$ distribution of the particle of interest will be centered around 0 in this new variable. Figure~\ref{fig:pidpion_raw} shows the typical PID lines zoomed in on the pions. Figure~\ref{fig:pidpion} shows the corresponding $z_i$ distribution after all the pitch-yaw bins have been merged. Both charged pions lie on $z_i=0$, where other particle species rapidly diverge.

\begin{figure}[!htb]
    \centering
    \begin{subfigure}[t]{0.49\textwidth}
        \centering
        \includegraphics[width=\linewidth]{pidPion.png}
        \caption{Output PID pions.} \label{fig:pidpion_raw}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{.49\textwidth}
        \centering
        \includegraphics[width=\linewidth]{pidPion_zi.png} 
        \caption{PID plot for $\pi^-$ and $\pi^+$ species in $z_i$. } \label{fig:pipion_zi}
    \end{subfigure}

    \caption{$\langle dE/dx\rangle$ versus p/q for $\tin{132}{124}$ system.}
\label{fig:pidpion}
\end{figure}

Now that the pion distribution is flattened, we take bin slices along $p/q$ to determine the particle yield in a certain bin and also estimate the background contribution. The background relevant to  the charged pions comes from the $e^+$ and $e^-$ resulting from the $\pi^0$ Dalitz decay described in Appendix~\ref{appen:dalitz}. Figure~\ref{fig:pimmom_flat} shows the Gaussian fits to the $z_i$ distribution for two different $p/q$ binned cuts around the $\pi^-$ particles, where as Fig.~\ref{fig:pipmom_flat} shows the background $\pi^+$ particles. A Gaussian fit is performed to the pions and to the $e$ spectrum to estimate the background contribution.

\begin{figure}[!htb]
    \centering
    \begin{subfigure}[t]{0.49\textwidth}
        \centering
        \includegraphics[width=\linewidth]{pim_mom2.png}
        \caption{Bin -54 MeV/c < p/q < -52. } \label{fig:pimmom2}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{.49\textwidth}
        \centering
        \includegraphics[width=\linewidth]{pim_mom1.png} 
        \caption{Bin -68 MeV/c < p/q < -66.} \label{fig:pimmom1}
    \end{subfigure}

    \caption{$z_i$ projection for binned slices in p/q around the $\pi^-$.}
\label{fig:pimmom_flat}
\end{figure}

\begin{figure}[!htb]
    \centering
    \begin{subfigure}[t]{0.49\textwidth}
        \centering
        \includegraphics[width=\linewidth]{pip_mom1.png}
        \caption{Bin 74 MeV/c < p/q < 76. } \label{fig:pipmom1}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{.49\textwidth}
        \centering
        \includegraphics[width=\linewidth]{pip_mom2.png} 
        \caption{Bin 94 MeV/c < p/q < 86.} \label{fig:pipmom2}
    \end{subfigure}

    \caption{$z_i$ projection for binned slices in p/q around the $\pi^+$.}
\label{fig:pipmom_flat}
\end{figure}

### 2020-06-14 chapters/ch_data_analysis_2.tex
Backup

\begin{table}[H]
  \begin{center}
    \begin{tabular}{cccccc}
      \hline 
      Particle Type & $\mu_\mathrm{A/Q}$ & $\sigma_\mathrm{A/Q}$ & $\mu_\mathrm{Z}$ &
      $\sigma_\mathrm{Z}$ & $\rho$\\
      \hline\hline 
      ${}^{132}$Sn & 2.64 & 0.0014 & 49.95 & 0.209 & -0.052 \\
    %  $\tin{124}{112}$ & 2.24005 & 0.00150934 & 49.9906 & 0.194804 & -0.0671999 \\
    %  $\tin{112}{124}$ & 2.47995 & 0.00182327 & 49.961 & 0.214745 & -0.0742297 \\
      ${}^{108}$Sn & 2.16 & 0.0015 & 49.99 & 0.207 & -0.059 \\
      \hline
    \end{tabular}
    \caption{Multivariate normal distribution fit parameters for four beams.
    \label{beamParameters}}
  \end{center}
\end{table}
Figure~\ref{fig:beampid132} shows a zoomed in view of the PID centered around the ${}^{132}$Sn beam and Fig.~\ref{fig:beampid108} the ${}^{108}$Sn beam. The red lines represent the cut where particles identified inside the circle represent the beam events which are identified as the good beam events. For both ${}^{132}$Sn an ${}^{108}$Sn a 2.83$\sigma$ cut is taken around the mean values.

@@

\section{Beam angle selection}
The incoming secondary beam is deflected by the magnetic field and impinges on the target at some small but significant angle. From the BDCs tracking information, the beam is projected as a straight line right up until entering the magnet.  The beam is then propagated through the magnetic field using a Runge-Kutta integration until reaching the target position. The beam angle on target can be categorized by two angles $\theta_{a_{proj}}$ and $\theta_{b_{proj}}$ defined as,

@@

\begin{table}[H]
  \begin{center}
    \begin{tabular}{ccccc}
      \hline 
      System & $\mu_{\theta_\mathrm{a,proj}}$ &
      $\sigma_{\theta_\mathrm{a,proj}}$ & $\mu_{\theta_\mathrm{b,proj}}$ &
      $\sigma_{\theta_\mathrm{b,proj}}$ \\
      \hline\hline 
      $\tin{132}{124}$ & 0.61 & 2.94 & -44.18 & 1.96 \\
    %  $\tin{112}{124}$ & -0.41 & 1.58 & -53.52 & 0.90 \\
      $\tin{108}{112}$ & -0.42 & 1.95 & -55.17 & 0.97 \\
      \hline
    \end{tabular}
    \caption{2D Gaussian fit parameters for $^{132}$Sn, $^{112}$Sn and
      $^{108}$Sn beam angle. \label{beamAngleParameters}}
  \end{center}
\end{table}

>>>>

\begin{table}[!htb]
  \begin{center}
    \begin{tabular}{cccccc}
      \hline 
      Particle Type & $\mu_\mathrm{A/Q}$ & $\sigma_\mathrm{A/Q}$ & $\mu_\mathrm{Z}$ &
      $\sigma_\mathrm{Z}$ & $\rho$\\
      \hline\hline 
      ${}^{132}$Sn & 2.64 & 0.0014 & 49.95 & 0.209 & -0.052 \\
    %  $\tin{124}{112}$ & 2.24005 & 0.00150934 & 49.9906 & 0.194804 & -0.0671999 \\
    %  $\tin{112}{124}$ & 2.47995 & 0.00182327 & 49.961 & 0.214745 & -0.0742297 \\
      ${}^{108}$Sn & 2.16 & 0.0015 & 49.99 & 0.207 & -0.059 \\
      \hline
    \end{tabular}
    \caption{Multivariate normal distribution fit parameters for four beams.
    \label{beamParameters}}
  \end{center}
\end{table}
Figure~\ref{fig:beampid132} shows a zoomed in view of the PID centered around the ${}^{132}$Sn beam and Fig.~\ref{fig:beampid108} the ${}^{108}$Sn beam. The red lines represent the cut where particles identified inside the circle represent the beam events which are identified as the good beam events. For both ${}^{132}$Sn an ${}^{108}$Sn a 2.83$\sigma$ cut is taken around the mean values.

@@

\section{Beam angle selection}
\label{sec:beamangle}
The incoming secondary beam is deflected by the magnetic field and impinges on the target at some small but significant angle. From the BDCs tracking information, the beam is projected as a straight line right up until entering the magnet.  The beam is then propagated through the magnetic field using a Runge-Kutta integration until reaching the target position. The beam angle on target can be categorized by two angles $\theta_{a_{proj}}$ and $\theta_{b_{proj}}$ defined as,

@@

\begin{table}[!htb]
  \begin{center}
    \begin{tabular}{ccccc}
      \hline 
      System & $\mu_{\theta_\mathrm{a,proj}}$ &
      $\sigma_{\theta_\mathrm{a,proj}}$ & $\mu_{\theta_\mathrm{b,proj}}$ &
      $\sigma_{\theta_\mathrm{b,proj}}$ \\
      \hline\hline 
      $\tin{132}{124}$ & 0.61 & 2.94 & -44.18 & 1.96 \\
    %  $\tin{112}{124}$ & -0.41 & 1.58 & -53.52 & 0.90 \\
      $\tin{108}{112}$ & -0.42 & 1.95 & -55.17 & 0.97 \\
      \hline
    \end{tabular}
    \caption{2D Gaussian fit parameters for $^{132}$Sn, $^{112}$Sn and
      $^{108}$Sn beam angle. \label{beamAngleParameters}}
  \end{center}
\end{table}

### 2020-06-14 chapters/ch_results.tex
Backup

The efficiency correction is applied track-by-track, where the efficiency of the i-th track $\epsilon_i$ is retrieved from the database. The correction factor $C_i$ is defined as,

@@

The track is then weighted by the factor $C_i$ when plotting in any type of graph for any observable.

%need to add error analysis

%Efficiency correction
%space charge before and after
%left and right of space charge 
%error analysis
%systematic error analysis

@@

@@

\section{$\pi^-$/$\pi^+$ Ratio}

>>>>

The efficiency correction is applied track-by-track, where the efficiency of the i-th track $\epsilon_i$ is retrieved from the database, from the parameters discussed in Section~\ref{sec:efficiency}. The correction factor $C_i$ is defined as,

@@

Each track that is identified as a pion, is then weighted by the correction factor $C_i$, when filling any histogram. By using the track-by-track method, we can transform the track into any observable of interest, notably we can transform the  track from the lab frame into the center-of-mass (CM) frame. Before performing the transformation to the CM system recall the beam is at a small, but non-negligible, angle in the Lab frame, see Section~\ref{sec:beamangle}. The beam angle for each even is measured and a rotation is applied to all the tracks in an event, to align the beam angle along the z-axis. Doing so makes the transformation into the CM system much simpler to describe. If the beam direction is defined by a unit vector $\hat{b}$, we can define the rotation that rotates the beam into the z-axis as a rotation about an arbitrary vector $\hat{v} = \hat{b}\times\hat{z}$ where the angle between the two is given by $\cos \theta = \hat{b}\cdot\hat{z}$.

Once all the events have been rotated to align with the z-axis, transforming from the Lab to the CM frame is done by a Lorentz transformation. Where the 4-momentum vector in the lab frame is defined as $\textbf{P} = (E/c,p_x,p_y,p_z)$. Where the corresponding Lorentz transform into the CM frame along the beam (z-axis) is defined as,

\begin{equation}
A = \begin{pmatrix}
1 & 0 & 0 & 0\\
0 & 1 & 0 & 0\\
0 & 0 & \gamma & -\beta \gamma\\
0 & 0 & -\beta \gamma & \gamma
\end{pmatrix},
\end{equation}

where $\beta$, describes the velocity of the CM system, and $\gamma=\sqrt{1-\beta^2}^{-1}$. The parameter $\beta$ can be determined from the total momentum of the system in the Laboratory frame $P = \sqrt{ T_{P}^2 - M_{T}^2}$ and the total energy of the system $E = T_{P} + M_{P} + M_{T}$ as $\beta = -P/E$, where the (-) sign denotes the correct direction for transforming from the Lab to the CM frame. The CM transformed track is defined as $p^{CM} = \textbf{A}p^{Lab}$.

%need to add error analysis
%systematic error analysis

Figure~\ref{fig:pionspectra} shows the pion CM kinetic energy spectra for both the $\tin{132}{124}$ and $\tin{108}{112}$ systems. This data marks the first pion spectral data at sub-threshold energies, extending to very low pion energies. The coulomb potential works to accelerate $\pi^+$ and decelerate $\pi^-$ particles due to the positive charge of the nuclear medium. The coulomb barrier prevents the production of low energy $\pi^+$, which is why there is a reduction in the phase space around zero kinetic energy.

@@

The integrated pion yield for both systems and the $\pi^-/\pi^+$ ratio is listed in Table~\ref{tb:pionyield}, where the systematic errors are the first error bar and the statistical error is listed next. It is remarkable that the pion ratio is significantly greater than the N/Z of the system which is what is expected in the Delta Resonance model or under the assumption of chemical equilibrium which is $\pi^-/\pi^+ = (N/Z)^2$ CITE HERE; 2x in the $\tin{132}{124}$ system and 1.4x in the $\tin{108}{112}$ system. The pion ratio was hypothesized to be proportional to the high density N/Z ratio of the early system, where the other effects such as pion absorption and re-emission would dilute the effect and lower the ratio as the system tends toward isospin equilibrium. A na\"ive interpretation would be the high density N/Z fraction is much higher for some reason than the two systems, though, the system spends very little time in this state which would require a very soft Symmetry Energy. A more likely conclusion is the mechanisms involved in pion productions via the $\Delta(1232)$ resonance is more complicated than the simple $(N/Z)^2$ relation. For example the $\Delta$ resonance potential has an iso-scalar component (independent of iso-spin) and an iso-vector component. The nature of these two potentials is still unknown CITE HERE, and the role it plays in pion production has shown to be very important CITE HERE.

@@

We have compared the total pion yields and ratios to the 6 common transport codes for the systems measured. The table of the transport codes are listed in Appendix~\ref{tb:pionyieldTheory}. Figure~\ref{fig:totalpiYield} shows the total pion yield for the four systems measured as compared with the codes. The codes plotted here are only the soft symmetry energy since the variation in code is much larger than the intra-code variation between different Symmetry Energies. While some codes make a reasonable approximation of a particular charge pion, no code reasonably predict both.

\begin{figure}[!htb]
\centering
\includegraphics[width=.8\textwidth]{totalpiYield.png}
\caption{Total pion yields as compared with 6 common transport codes.}
\label{fig:totalpiYield}
\end{figure}

Figure~\ref{fig:totalpiRatio} shows the total single pion ratio and the double ratio of the $\tin{132}{124}$ and $\tin{108}{112}$ system. Here, the variation between codes is much larger than the variation between the Symmetry Energy within a particular code. The Symmetry Energy variation of two codes -- $\chi$BUU and TuQMD -- is plotted as a wide band in the single ratio and all codes in the double ratio; the data band represents the data error bar. Certainly the variation between Symmetry Energy in the $\tin{132}{124}$, though small, still exists in many of the codes as predicted initially CITE HERE BAO ANN. The small error bars in the data also would allow for a detailed analysis to extract a constraint of the Symmetry Energy, if the variation between codes could be solved.

\begin{figure}[!htb]
\centering
\includegraphics[width=\textwidth]{totalpiRatio.png}
\caption{Total pion ratio and double ratio compared with 6 common transport codes.}
\label{fig:totalpiRatio}
\end{figure}

\section{Pion Spectral Ratio}
